Expectations, Outcomes, and Challenges \\ of Modern Code Review

### 2013-03-01 sections/conclusion.tex
did another full pass on the paper

In this work, we investigated modern, tool-based code review, uncovered both a wide range of motivations for review, and determined that the outcomes do not always match those motivations. We identified understanding as a key component and provided recommendations to both practitioners and researchers. It is our hope that the insights we have discovered lead to more effective review in practice and improved tools, based on research, to aid developers perform code reviews.

>>>>

We investigated modern, tool-based code review, uncovering both a wide range of motivations for review and that the outcomes do not always match those motivations. We identified understanding as a key component and provided recommendations to both practitioners and researchers. It is our hope that the insights we have discovered lead to more effective review in practice and improved tools, based on research, to aid developers perform code reviews.

### 2013-03-01 sections/introduction.tex
did another full pass on the paper

Nowadays many organizations are adopting more lightweight code review
practices to limit the inefficiencies of inspections. In particular, there is a
clear trend toward the usage of tools developed to support code
review~\cite{rigby2012open}. In the context of this paper, we define
\emph{Modern Code Review}, as review that is \begin{inparaenum}[(1)]
\item informal (in contrast to Fagan-style), 
\item tool-based, and that 
\item occurs regularly in practice nowadays, for example at companies such as Microsoft, Google~\cite{kennedy2006online},
Facebook~\cite{tsotsis2011online}, and in other organizations and OSS projects~\cite{gerrit2012online}. \end{inparaenum}

@@

Answers to these questions can provide insight for both practitioners and
researchers.  Developers and other software project stakeholders can use
empirical evidence about expectations and outcomes to make informed decisions
about when to use code review and how it should fit into their development
process. Researchers can focus their attention on practitioners' challenges to
make code review more effective.

@@

Our results show that, although the top motivation driving code reviews is
still finding defects, the practice and the actual outcomes are less about
finding errors than expected: Defect related comments comprise a small
proportion and mainly cover small logical low-level issues. At the same time,
code review additionally provides a wide spectrum of benefits to software
teams, such as knowledge transfer, team awareness, and improved solutions to
problems. Moreover, we found that context and change understanding is the key
of any review. According to the outcomes they want to achieve, developers
employ many mechanisms to fulfill their understanding needs, most of which are
not currently met by any code review tool.

>>>>

Nowadays many organizations are adopting more lightweight code review
practices to limit the inefficiencies of inspections. In particular, there is a
clear trend toward the usage of tools developed to support code
review~\cite{rigby2012open}. In the context of this paper, we define
\emph{Modern Code Review}, as review that is \begin{inparaenum}[(1)]
\item informal (in contrast to Fagan-style), 
\item tool-based, and that 
\item occurs regularly in practice nowadays, for example at companies such as Microsoft, Google~\cite{kennedy2006online},
Facebook~\cite{tsotsis2011online}, and in other organizations and open source software (OSS) projects~\cite{gerrit2012online}. \end{inparaenum}

@@

Answers to these questions can provide insight for both practitioners and
researchers.  Developers and other software project stakeholders can use
empirical evidence about expectations and outcomes to make informed decisions
about when to use code review and how it should fit into their development
process. Researchers can focus their attention on the challenges faced by practitioners to make code review more effective.

@@

Our results show that, although the top motivation driving code reviews is
finding defects, the practice and the actual outcomes are less about
finding errors than expected: Defect related comments comprise a small
proportion and mainly cover small logical low-level issues. On the other hand,
code review additionally provides a wide spectrum of benefits to software
teams, such as knowledge transfer, team awareness, and improved solutions to
problems. Moreover, we found that context and change understanding is the key
of any review. According to the outcomes they want to achieve, developers
employ many mechanisms to fulfill their understanding needs, most of which are
not currently met by any code review tool.

### 2013-03-01 sections/methodology.tex
added reference to tech report for seeing example card
did another full pass on the paper

Our investigation of code review revolves around the following research
questions, which we iteratively refined during our initial in-field
observations and interviews:

@@

Once reviewers open a CodeFlow review, they interact with it via a single
desktop window (\figref{fig:codeflow:screenshot}). On the top left (\textbf{1}), they see the list of files
changed in the current submission, plus a ``description.txt'' file, which
contains a textual explanation of the change, written by the author. On bottom
left, CodeFlow shows the list of reviewers and their status (\textbf{2}). We see that
Christian is the review author and Alberto, Tom, and Nachi are the reviewers.
Alberto has reviewed and is waiting for the author to act, as the clock icon
suggests, while Nachi already signed off on the changes. CodeFlow's main view
(\textbf{3}) shows the diff-highlighted content of the file currently under review. Both
the reviewers and the author can highlight portions of the code and add
comments inline (\textbf{4}). These comments can start threads of discussion and are the
interaction points for the people involved in the review. Each user viewing the
same review in CodeFlow sees events as they happen.  Thus, if an author and
reviewer are working on the review at the same time, the communication is
synchronous and comment threads act similar to instant messaging. The comments
are persisted so that if they work at different times, the communication
becomes asynchronous. The bottom right pane (\textbf{5}) shows the summary of all the
comments in the review.

CodeFlow centralizes and records all the information on code reviews on a
central server. This provides an additional data source that we used to analyze
real code review comments without incurring the Hawthorne effect~\cite{adair1984hawthorne}.

@@

Our research method followed a mixed qualitative and quantitative approach~\cite{creswell2009research}, depicted in
\figref{fig:research-method}, collecting data from different sources for triangulation: (\textbf{1})
analysis of previous study, (\textbf{2}) observations and interviews with developers,
(\textbf{3}) card sort on interview data,  (\textbf{4}) card sort on code review comments, (\textbf{5})
the creation of an affinity diagram, and (\textbf{6}) survey to managers and
programmers.

@@

Most of the interview questions revolved around topics that are very specific
to tool usage, and were only tangentially related to this work. We found one
relevant as a starting point for our study: ``\emph{What do you hope to accomplish
when you submit a code review?}'' We analyzed the transcript of this answer, for
each interview, through the process of \emph{coding}~\cite{berg2004qualitative} (also used in
\emph{grounded theory}~\cite{adolph2011using}): breaking up the answers into smaller coherent
units (sentences or paragraphs) and adding \emph{codes} to them. We organized codes
into \emph{concepts}, which in turn were grouped into more abstract \emph{categories}.

@@

\textbf{2. Observations and interviews with developers:} Next, we conducted a series of
one-to-one meetings with developers who use CodeFlow, each taking 40-60
minutes.

@@

The second part of the meeting was a \emph{semi-structured} interview~\cite{taylor2010qualitative}. Semi-structured interviews make use of an \emph{interview guide} that
contains general groupings of topics and questions rather than a pre-determined
exact set and order of questions.  They are often used in an exploratory
context to ``find out what is happening [and] to seek new insights''~\cite{weiss1995learning}. 
The guideline was iteratively refined after each interview, in
particular when developers started providing answers very similar to the
earlier ones, thus reaching a saturation effect.

After the first 5-6 meetings, the observations reached a \emph{saturation} point~\cite{Glas1998a}: They were providing insights very similar to the earlier ones. For this, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a ``hook'' to talk about topics in our guideline.

@@

The first author of this paper created all of the cards, from the 1,047
coherent units in the interviews. Throughout our further analysis other
researchers (the second author and external people) were involved in developing
categories and assigning cards to categories, so as to strengthen the validity
of the result. The first author played a special role of ensuring that the
context of each question was appropriately considered in the categorization,
and creating the initial categories. To ensure the integrity of our categories,
the cards were sorted by the first author several times to identify initial
themes. To reduce bias from the first author sorting the cards to form initial themes, all researchers reviewed and agreed on the final set of
categories.

\textbf{4. Card sort (code review comments):} The same method was applied to group code
review comments into categories: We randomly sampled 200 threads with at least
two comments (e.g., Point 4 of \figref{fig:research-method}), from the entire dataset of CodeFlow
reviews, which embeds data from dozens of independent software products at
Microsoft. We printed one card for each comment (along with the entire
discussion thread to give the context), totaling 570 cards, and conducted a
card sort, as performed for the interviews, to identify common themes.

@@

\textbf{6. Surveys:} The final step of our study was aimed at validating the
concepts that emerged from the previous phases. Towards this goal, we created
two surveys to reach a significant number of participants and to challenge our
conclusions (The full surveys are available as a technical report~\cite{bacchelli2012appendix}). 
For the design of the surveys, we followed Kitchenham and
Pfleeger's guidelines for personal opinion surveys~\cite{kitchenham2008personal}. 
Both surveys were anonymous to increase response rates~\cite{tyagi1989effects}.

>>>>

Our investigation of code review revolves around the following research
questions (iteratively refined during our initial in-field
observations and interviews):

@@

Once reviewers open a CodeFlow review, they interact with it through a single
desktop window (\figref{fig:codeflow:screenshot}). On the top left (\textbf{1}), they see the list of files
changed in the current submission, plus a `description.txt' file, which
contains a textual explanation of the change, written by the author. On bottom
left, CodeFlow shows the list of reviewers and their status (\textbf{2}). We see that
Christian is the review author and Alberto, Tom, and Nachi are the reviewers.
Alberto has reviewed and is waiting for the author to act, as the clock icon
suggests, while Nachi already signed off on the changes. CodeFlow's main view
(\textbf{3}) shows the diff-highlighted content of the file currently under review. Both
the reviewers and the author can highlight portions of the code and add
comments inline (\textbf{4}). These comments can start threads of discussion and are the
interaction points for the people involved in the review. Each user viewing the
same review in CodeFlow sees events as they happen.  Thus, if an author and
reviewer are working on the review at the same time, the communication is
synchronous and comment threads act similar to instant messaging. The comments
are persisted so that if they work at different times, the communication
becomes asynchronous. The bottom right pane (\textbf{5}) shows the summary of all the
comments in the review.

CodeFlow records all the information on code reviews on a
central server. This provides an additional data source that we used to analyze
real code review comments without incurring the Hawthorne effect~\cite{adair1984hawthorne}.

@@

Our research method followed a mixed qualitative and quantitative approach~\cite{creswell2009research} 
(depicted in \figref{fig:research-method}), which collects data from different sources for triangulation: (\textbf{1})
analysis of previous study, (\textbf{2}) observations and interviews with developers,
(\textbf{3}) card sort on interview data,  (\textbf{4}) card sort on code review comments, (\textbf{5})
the creation of an affinity diagram, and (\textbf{6}) survey to managers and
programmers.

@@

Most of the interview questions revolved around topics that were very specific
to usage of the tool, and were only tangentially related to this work. We found one
relevant as a starting point for our study: ``\emph{What do you hope to accomplish
when you submit a code review?}'' We analyzed the transcript of this answer, for
each interview, through the process of \emph{coding}~\cite{berg2004qualitative} (also used in
\emph{grounded theory}~\cite{adolph2011using}): breaking up the answers into smaller coherent
units (sentences or paragraphs) and adding \emph{codes} to them. We organized codes
into \emph{concepts}, which in turn were grouped into more abstract \emph{categories}.

@@

\textbf{2. Observations and interviews with developers:} Subsequently, we conducted a series of
one-to-one meetings with developers who use CodeFlow, each taking 40-60
minutes.

@@

The second part of the meeting was a \emph{semi-structured} interview~\cite{taylor2010qualitative}. This form of interviews makes use of an \emph{interview guide} that
contains general groupings of topics and questions rather than a pre-determined
exact set and order of questions.  They are often used in an exploratory
context to \quotation{find out what is happening \emph{[and]} to seek new insights}~\cite{weiss1995learning}. 
The guideline was iteratively refined after each interview, in
particular when developers started providing answers very similar to the
earlier ones, thus reaching a saturation effect.

After the first 5-6 meetings, the observations reached a \emph{saturation} point~\cite{Glas1998a}: They were providing insights very similar to the earlier ones. For this, we adjusted the meetings to have shorter observations, which we used as a starting point for our meetings and as a hook to talk about topics in our guideline.

@@

The first author of this paper created all of the cards, from the 1,047
coherent units in the interviews (an example card is shown in a technical report~\cite{bacchelli2012appendix}). Throughout our further analysis other
researchers (the second author and external people) were involved in developing
categories and assigning cards to categories, so as to strengthen the validity
of the result. The first author played a special role of ensuring that the
context of each question was appropriately considered in the categorization,
and creating the initial categories. To ensure the integrity of our categories,
the cards were sorted by the first author several times to identify initial
themes. To reduce bias from the first author sorting the cards to form initial themes, all researchers reviewed and agreed on the final set of
categories.

\textbf{4. Card sort (code review comments):} The same method was applied to group code
review comments into categories: We randomly sampled 200 threads with at least
two comments (\eg Point 4 of \figref{fig:research-method}), from the entire dataset of CodeFlow
reviews, which embeds data from dozens of independent software products at
Microsoft. We printed one card for each comment (along with the entire
discussion thread to give the context), totaling 570 cards, and conducted a
card sort, as performed for the interviews, to identify common themes.

@@

\textbf{6. Surveys:} The final step of our study was aimed at validating the
concepts that emerged from the previous phases. Towards this goal, we created
two surveys to reach a significant number of participants and to challenge our
conclusions (the full surveys are available as a technical report~\cite{bacchelli2012appendix}). 
For the design of the surveys, we followed Kitchenham and
Pfleeger's guidelines for personal opinion surveys~\cite{kitchenham2008personal}. 
Both surveys were anonymous to increase response rates~\cite{tyagi1989effects}.

### 2013-03-01 sections/related_work.tex
did another full pass on the paper

Previous studies exist that have examined the practices of code inspection and
code review.  Stein \etal conducted a study focusing specifically on
distributed, asynchronous code inspections~\cite{stein1997casestudy}. The study included evaluation
of a tool that allowed for identification and sharing of code faults or
defects. Participants at separated locations can then discuss faults via the
tool. Laitenburger conducted a survey of code inspection methods, and presented
a taxonomy of code inspection techniques~\cite{laitenberger2002survey}. Johnson conducted
an investigation into code review practices in open source development and their
effect on choices made by software project managers~\cite{johnson2006collaboration}.

@@

More recently, Rigby has done extensive work examining code review practices in
open source software development~\cite{rigby2012open}.  For example in
a study of practices in the Apache project~\cite{rigby2008open}
they data-mined the email archives and found that reviews were typically small
and frequent, and that the contributions to a review were often brief and
independent from one another.

Sutherland and Venolia conducted a study at Microsoft regarding using code
review data for later information needs~\cite{sutherland2009can}. They
hypothesized that the knowledge exchanged during code reviews could be of great
value to engineers later trying to understand or modify the discussed code.
They found that ``the meat of the code review dialog, no matter what medium, is
the articulation of design rationale'' and, thus, ``code reviews are an enticing
opportunity for capturing design rationale.''

>>>>

Previous studies have examined the practices of code inspection and
code review.  Stein \etal conducted a study focusing specifically on
distributed, asynchronous code inspections~\cite{stein1997casestudy}. The study included evaluation
of a tool that allowed for identification and sharing of code faults or
defects. Participants at separated locations can then discuss faults via the
tool. Laitenburger conducted a survey of code inspection methods, and presented
a taxonomy of code inspection techniques~\cite{laitenberger2002survey}. Johnson conducted
an investigation into code review practices in OSS development and their
effect on choices made by software project managers~\cite{johnson2006collaboration}.

@@

More recently, Rigby has done extensive work examining code review practices in
OSS development~\cite{rigby2008open, Rigb2011a, rigby2012open}.  For example in
a study of practices in the Apache project~\cite{rigby2008open},
they data-mined the email archives and found that reviews were typically small
and frequent, and that the contributions to a review were often brief and
independent from one another.

Sutherland and Venolia conducted a study at Microsoft regarding using code
review data for later information needs~\cite{sutherland2009can}. They
hypothesized that the knowledge exchanged during code reviews could be of great
value to engineers later trying to understand or modify the discussed code.
They found that \quotation{the meat of the code review dialog, no matter what medium, is
the articulation of design rationale} and, thus, \quotation{code reviews are an enticing
opportunity for capturing design rationale.}
